Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-08-06 submission/manuscript.Rmd
final results update based on Ada suggesionts

**The most important OTUs in Random Forest models built to classify SCFA concentrations or tumors are not the same.** It is possible that due to the way the Random Forest algorithm works, OTUs associated with SCFA concentrations could be downweighted in importance within the adenoma or carcinoma models when SCFA concentrations are included. To test if this is the case we used OTU data and built Random Forest models to classify SCFA concentrations. Although the correlation between the predicted and actual SCFA concentrations were moderately associated with each other [Figure 5A], there was a difference in accuracy based on whether the fecal sample was from an individual without tumors or from patients with adenomas or carcinomas [Figure 5B]. Additionally, because the training set R^2^ was always higher than the test set R^2^, all SCFA concentration models tended to be over fit, suggesting that rarer taxa are important for these classifications [Figure 5A]. When comparing the adenoma or carcinoma model to the SCFA concentration models there was minimal overlap between these model's most important OTUs [Figure 4B-C, 4E-F and 5C-E]. The only OTU that was in the top 10 most important variables that had overlap between the models was OTU00167 (*Clostridiales*) [Figure 4B-C, 4E-F, 5C-E]. Similar observations were made when using high/low SCFA groups based on higher/lower than median SCFA concentrations [Figure S2]. Collectively, these observations provide evidence that it is possible to identify specific OTUs associated with higher SCFA concentrations and accordingly these OTUs belong to taxa known to produce acetate, butyrate, and propionate. Although it is possible to identify OTUs associated with SCFA production, overall, our results do not support the hypothesis that SCFA concentration or OTUs associated with their production are different between individuals with no tumors and patients with adenomas or carcinomas.

>>>>

**The most important OTUs in Random Forest models built to classify SCFA concentrations or tumors are not the same.** It is possible that due to the way the Random Forest algorithm works, OTUs associated with SCFA concentrations could be downweighted in importance within the adenoma or carcinoma models when SCFA concentrations are included. To test if this is the case we used OTU data and built Random Forest models to classify SCFA concentrations. Overall, the correlation between the predicted and actual SCFA concentrations were moderately associated with each other [Figure 5A]. Additionally, because the training set R^2^ was always higher than the test set R^2^, all SCFA concentration models tended to be over fit, suggesting that rarer taxa were important for these classifications [Figure 5A]. There also was a difference in accuracy based on whether the fecal sample was from an individual without tumors or from patients with adenomas or carcinomas [Figure 5B]. When comparing the adenoma or carcinoma model to the SCFA concentration models there was minimal overlap between these model's most important OTUs [Figure 4B-C, 4E-F and 5C-E]. The only OTU that was in the top 10 most important variables and had overlap between the models was OTU00167 (*Clostridiales*) [Figure 4B-C, 4E-F, 5C-E]. Similar observations were made when using high/low SCFA groups based on the median SCFA concentration [Figure S2]. Collectively, these observations provide evidence that it is possible to identify specific OTUs associated with higher SCFA concentrations and accordingly these OTUs belong to taxa known to produce acetate, butyrate, and propionate. Although it is possible to identify OTUs associated with SCFA production, overall, our results do not support the hypothesis that SCFA concentration or OTUs associated with their production are different between individuals with no tumors and patients with adenomas or carcinomas.

### 2018-08-06 submission/manuscript.tex
final results update based on Ada suggesionts

\textbf{The most important OTUs in Random Forest models built to
classify SCFA concentrations or tumors are not the same.} It is possible
that due to the way the Random Forest algorithm works, OTUs associated
with SCFA concentrations could be downweighted in importance within the
adenoma or carcinoma models when SCFA concentrations are included. To
test if this is the case we used OTU data and built Random Forest models
to classify SCFA concentrations. Although the correlation between the
predicted and actual SCFA concentrations were moderately associated with
each other {[}Figure 5A{]}, there was a difference in accuracy based on
whether the fecal sample was from an individual without tumors or from
patients with adenomas or carcinomas {[}Figure 5B{]}. Additionally,
because the training set R\textsuperscript{2} was always higher than the
test set R\textsuperscript{2}, all SCFA concentration models tended to
be over fit, suggesting that rarer taxa are important for these
classifications {[}Figure 5A{]}. When comparing the adenoma or carcinoma
model to the SCFA concentration models there was minimal overlap between
these model's most important OTUs {[}Figure 4B-C, 4E-F and 5C-E{]}. The
only OTU that was in the top 10 most important variables that had
overlap between the models was OTU00167 (\emph{Clostridiales}) {[}Figure
4B-C, 4E-F, 5C-E{]}. Similar observations were made when using high/low
SCFA groups based on higher/lower than median SCFA concentrations
{[}Figure S2{]}. Collectively, these observations provide evidence that
it is possible to identify specific OTUs associated with higher SCFA
concentrations and accordingly these OTUs belong to taxa known to
produce acetate, butyrate, and propionate. Although it is possible to
identify OTUs associated with SCFA production, overall, our results do
not support the hypothesis that SCFA concentration or OTUs associated
with their production are different between individuals with no tumors
and patients with adenomas or carcinomas.

>>>>

\textbf{The most important OTUs in Random Forest models built to
classify SCFA concentrations or tumors are not the same.} It is possible
that due to the way the Random Forest algorithm works, OTUs associated
with SCFA concentrations could be downweighted in importance within the
adenoma or carcinoma models when SCFA concentrations are included. To
test if this is the case we used OTU data and built Random Forest models
to classify SCFA concentrations. Overall, the correlation between the
predicted and actual SCFA concentrations were moderately associated with
each other {[}Figure 5A{]}. Additionally, because the training set
R\textsuperscript{2} was always higher than the test set
R\textsuperscript{2}, all SCFA concentration models tended to be over
fit, suggesting that rarer taxa were important for these classifications
{[}Figure 5A{]}. There also was a difference in accuracy based on
whether the fecal sample was from an individual without tumors or from
patients with adenomas or carcinomas {[}Figure 5B{]}. When comparing the
adenoma or carcinoma model to the SCFA concentration models there was
minimal overlap between these model's most important OTUs {[}Figure
4B-C, 4E-F and 5C-E{]}. The only OTU that was in the top 10 most
important variables and had overlap between the models was OTU00167
(\emph{Clostridiales}) {[}Figure 4B-C, 4E-F, 5C-E{]}. Similar
observations were made when using high/low SCFA groups based on the
median SCFA concentration {[}Figure S2{]}. Collectively, these
observations provide evidence that it is possible to identify specific
OTUs associated with higher SCFA concentrations and accordingly these
OTUs belong to taxa known to produce acetate, butyrate, and propionate.
Although it is possible to identify OTUs associated with SCFA
production, overall, our results do not support the hypothesis that SCFA
concentration or OTUs associated with their production are different
between individuals with no tumors and patients with adenomas or
carcinomas.
